### 2022-04-15 1_page_request.tex
First commit, request finished, paper not began

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Proposition of New Experiment to Better Understand the Relation Between Typicality and Prototypes%\\
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
%should not be used}
%\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Samuel Kostadinov}
%\IEEEauthorblockA{\textit{Department of } \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
}

\maketitle

\section{Introduction}
In this paper, I would like to propose an experiment to gain higher knowledge about the bound between typicality and key features of an object. 
The first step of the experiment is to extract the key features of an image. 
After extracting the features, we build a prototype of the image. 
As the last step, through a siamese network, we evaluate the similarity of an image. 
This process can help us understand if there is a bond between the typicality and resemblance to a prototype.

\section{The experiment}
The experiment I would like to suggest is composed of some phases. The first thing to do is to collect the images for the experiment. 
The collected images should be part of different categories like, for example, "birds" or "dogs". 
Then there is data collection about human typicality judgments. To do this, there should be a survey in which users rate the typicality of the previously collected images. 
Once the data are complete, in order to reduce possible noise, we should average the data.  
While collecting the data, there is the possibility to start working on the images. 
The first step is to extract meaningful features. 
Some techniques that range from simple edge detection to deformable shape analysis may be functional.
Other strategies that require the presence of neural networks, either fully connected or convolutional, are used in medical imaging and can be useful if adapted to a non-medical domain. 
Moreover, if the dataset contains hyperspectral images, some specific techniques that use convolutional or recurrent neural networks can be used to extract meaningful features. 
The following step is to build a prototype image based on the features extracted.
After building the prototype, the following step is to train a deep learning siamese network that will give a similarity estimation. 
This kind of network gives a similarity measure between two inputs. 
A siamese network's training requires a loss function that measures the distance between the two examples fed to the network. 
This can be helpful since if we use a custom loss function, defined based on human typicality rates, we can have an estimation on how similar is an image to the prototype that humans have. 
Moreover, if a different loss function is used, there is the possibility to understand how similar are the neural network's and the human's similarity concepts.

\section*{Some references}

\begin{itemize}

	\item Edge extraction methods \cite{b1}
	\item Shape correspondence \cite{b2}, \cite{b3}
	\item Medical imaging feature extraction \cite{b4}
	\item CNN feature extraction for pneumonia \cite{b5}
	\item Hybrid feature extraction for brain tumour classification \cite{b6}
	\item Hyperspectral feature extraction methods \cite {b7}
	%\item Siamese Networoks \cite{b8}
	\item Learning similarity metric for face verification \cite{b8}
	\item Learning similarity metrics (Cutting-Edge Trends in Deep Learning and Recognition course, University of Ilinois) \cite{b9}
	\item Understanding the contrastive loss reference \cite{b9}

\end{itemize}

\begin{thebibliography}{00}

\bibitem{b1} Owotogbe, JS and Ibiyemi, TS and Adu, BA, ``Edge detection techniques on digital images-a review'', Int J Innov Sci Res Technol, vol 2, pp. 329--332, 2019
\bibitem{b2}Sahillioğlu, Y. Recent advances in shape correspondence. Vis Comput 36, 1705–1721 (2020). https://doi.org/10.1007/s00371-019-01760-0
\bibitem{b3} Halimi, Oshri and Litany, Or and Rodola, Emanuele and Bronstein, Alex M and Kimmel, Ron, ``Unsupervised learning of dense shape correspondence'', Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition, pp. 4370--4379, 2019
\bibitem{b4} Chiranji Lal Chowdhary, D.P. Acharjya, ``Segmentation and Feature Extraction in Medical Imaging: A Systematic Review, Procedia Computer Science'', vol. 167, pp. 26-36, 2020, ISSN 1877-0509, https://doi.org/10.1016/j.procs.2020.03.179
\bibitem{b5} Varshni, Dimpy and Thakral, Kartik and Agarwal, Lucky and Nijhawan, Rahul and Mittal, Ankush, ``Pneumonia detection using CNN based feature extraction'', 2019 IEEE international conference on electrical, computer and communication technologies (ICECCT), pp. 1--7, 2019
\bibitem{b6} Gumaei, Abdu and Hassan, Mohammad Mehedi and Hassan, Md Rafiul and Alelaiwi, Abdulhameed and Fortino, Giancarlo, ``A hybrid feature extraction method with regularized extreme learning machine for brain tumor classification'', IEEE Access, vol. 7, pp. 36266--36273, 2019
\bibitem{b7} B. Rasti et al., "Feature Extraction for Hyperspectral Imagery: The Evolution From Shallow to Deep: Overview and Toolbox," in IEEE Geoscience and Remote Sensing Magazine, vol. 8, no. 4, pp. 60-88, Dec. 2020, doi: 10.1109/MGRS.2020.2979764
%\bibitem{b8} https://en.wikipedia.org/wiki/Siamese_neural_network
\bibitem{b8} Chopra, S.; Hadsell, R.; LeCun, Y., ``Learning a similarity metric discriminatively, with application to face verification'', 2005 IEEE Computer Society Conference on Computer Vision and Pattern Recognition, vol.1, pp. 539–546, doi:10.1109/CVPR.2005.202, 
\bibitem{b9} Chatterjee, Moitreya; Luo, Yunan, Similarity Learning with (or without) Convolutional Neural Network, 2017
\bibitem{b9} Wang, Feng and Liu, Huaping, ``Understanding the behaviour of contrastive loss'', Proceedings of the IEEE/CVF conference on computer vision and pattern recognition, pp. 2495--2504, 2021

%\bibitem{b10} Wang, Feng and Liu, Huaping; ``Understanding the behaviour of contrastive loss''; Proceedings of the IEEE/CVF conference on computer vision and pattern recognition; pp. 2495--2504, 2021

\end{thebibliography}
\vspace{12pt}

\end{document}